Concurrent Programming with CXL

### 2020-06-26 book/paper.tex
Working on pretty printing code
Working on formatting CXL code

\documentclass{report}
\usepackage{fullpage}
\usepackage{imakeidx}\makeindex
\usepackage{hyperref}
\usepackage[nottoc,notlot,notlof]{tocbibind}
\usepackage[toc]{glossaries}\makeglossaries
\usepackage{graphicx}
\usepackage{framed}
\usepackage{tcolorbox}
\usepackage{verbatim}
\usepackage{fancyvrb}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{geometry}
\usepackage{enumitem}
\usepackage{etoolbox}

@@

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/triangle.cxl}
\end{code}
\caption{[triangle.cxl] Computing triangle numbers.}
\label{fig:triangle}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/Up.cxl}
\end{code}
\caption{[Up.cxl] Incrementing twice in parallel.}
\label{fig:inc}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/UpEnter.cxl}
\end{code}
\caption{[UpEnter.cxl] Broken attempt at fixing the code of Figure~\ref{fig:inc}.}
\label{fig:incenter}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/csbarebones.cxl}
\end{code}
\caption{[csbarebones.cxl] A barebones critical section with two processes.}
\label{fig:csbarebones}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/cs.cxl}
\end{code}
\caption{[cx.cxl] CXL model of a critical section.}
\label{fig:cs}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/naiveLock.cxl}
\end{code}
\caption{[naiveLock.cxl] Na\"{\i}ve implementation of a shared lock.}
\label{fig:uplock}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/naiveFlags.cxl}
\end{code}
\caption{[naiveFlags.cxl] Na\"{\i}ve use of flags to solve mutual exclusion.}
\label{fig:upflags}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/naiveTurn.cxl}
\end{code}
\caption{[naiveTurn.cxl] Na\"{\i}ve use of turn variable to solve mutual exclusion.}
\label{fig:upturn}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/Peterson.cxl}
\end{code}
\caption{[Peterson.cxl] Peterson's Algorithm}
\label{fig:peterson}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/PetersonInductive.cxl}
\end{code}
\caption{[PetersonInductive.cxl] Peterson's Algorithm with Inductive Invariant}
\label{fig:petersonproof}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/csonebit.cxl}
\end{code}
\caption{[csonebit.cxl] Mutual exclusion using a flag per process.}
\label{fig:csonebit}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/PetersonMethod.cxl}
\end{code}
\caption{[PetersonMethod.cxl] Peterson's Algorithm accessed through methods.}
\label{fig:petersonmethods}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/spinlock.cxl}
\end{code}
\caption{[spinlock.cxl] Mutual Exclusion using a ``spinlock'' based on test-and-set.}
\label{fig:tas}
\end{figure}

@@

While the CVM does not have any specific built-in interlock instructions,
it does have support for executing multiple instructions atomically.
This feature is available in the CXL language in two ways.
First, any CXL statement can be made atomic by placing a label in front
of it.  Second, a group of CXL statements can be made atomic
through its \texttt{atomic}
\index{atomic statement}
statement.
We can use \texttt{atomic} statement blocks to implement a wide variety of
interlock operations.
For example, we could fix the program in Figure~\ref{fig:inc} by
constructing an atomic increment operation for a counter, like so:
\begin{code}
\begin{Verbatim}[xleftmargin=5mm,numbers=left]
def atomic_inc(ptr):
    atomic:
        ^ptr += 1;
    ;
;
count = 0;
atomic_inc(&count);
\end{Verbatim}
\end{code}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/spinlockInv.cxl}
\end{code}
\caption{[spinlockInv.cxl] Checking invariants.}
\label{fig:tasinv}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/csTAS.cxl}
\end{code}
\caption{[csTAS.cxl] Fixed version of Figure~\ref{fig:uplock} using test-and-set.}
\label{fig:tas2}
\end{figure}

@@

\begin{figure}
\begin{code}
\begin{Verbatim}[xleftmargin=5mm,numbers=left]
def tas(lk):
    atomic:
        result = ^lk;
        ^lk = True;
    ;
;
def Lock():
    result = False;
;
def lock(lk):
    while tas(lk):
        pass;
    ;
;
def unlock(lk):
    ^lk = False;
;
\end{Verbatim}
\end{code}
\caption{The \texttt{Lock} interface in the \texttt{synch} module.}
\label{fig:spinlocks}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/UpLock.cxl}
\end{code}
\caption{[UpLock.cxl] Program of Figure~\ref{fig:inc} fixed with a lock.}
\label{fig:incfixed}
\end{figure}

@@

\begin{figure}
\begin{code}
\begin{Verbatim}[xleftmargin=5mm,numbers=left]
import list;

def Lock():
    result = dict{ .locked: False, .suspended: [] };
;
def lock(lk):
    atomic:
        if (^lk).locked:
            stop (^lk).suspended;
            assert (^lk).locked;
        else:
            (^lk).locked = True;
        ;
    ;
;
def unlock(lk):
    atomic:
        if (^lk).suspended == []:
            (^lk).locked = False;
        else:
            go (head((^lk).suspended)) ();
            (^lk).suspended = tail((^lk).suspended);
        ;
    ;
;
\end{Verbatim}
\end{code}
\caption{The \texttt{Lock} interface in the \texttt{synchS} module uses suspension.}
\label{fig:suspension}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/RW.cxl}
\end{code}
\caption{[RW.cxl] Busy-Waiting Reader/Writer Lock implementation.}
\label{fig:rwbusy}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/RWtest.cxl}
\end{code}
\caption{[RWtest.cxl] Test code for Figure~\ref{fig:rwbusy}.}
\label{fig:rwbusytest}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/RWbusychk.cxl}
\end{code}
\caption{[RWbusychk.cxl] Checking for Busy Waiting.}
\label{fig:rwblock}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/RWlock.cxl}
\end{code}
\caption{[RWlock.cxl] Reader/Writer with Two Locks.}
\label{fig:rw2lock}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/RWmulti.cxl}
\end{code}
\caption{[RWmulti.cxl] Checking that multiple readers can acquire the read lock.}
\label{fig:rwmulti}
\end{figure}

@@

\begin{figure}
\begin{code}
\begin{Verbatim}[xleftmargin=5mm,numbers=left]
def Semaphore(cnt):
    result = cnt;
;
def P(sema):
    let blocked = True:
        while blocked:
            atomic:
                if (^sema) > 0:
                    ^sema -= 1;
                    blocked = False;
                ;
            ;
        ;
    ;
;
def V(sema):
    atomic:
        ^sema += 1;
    ;
;
\end{Verbatim}
\end{code}
\caption{The \texttt{Semaphore} interface in the \texttt{synch} module.}
\label{fig:semaphore}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/DinersSema.cxl}
\end{code}
\caption{[DinersSema.cxl] Italian Philosophers.}
\label{fig:dinerssema}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/qsort.cxl}
\end{code}
\caption{[qsort.cxl] Iterative qsort() implementation.}
\label{fig:qsort}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/qsorttest.cxl}
\end{code}
\caption{[qsorttest.cxl] Test program for Figure~\ref{fig:qsort}.}
\label{fig:qsorttest}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=2.5in]{figures/pc-crop.pdf}
\end{center}
\caption{High-level state diagram specification of producers/consumers.}
\label{fig:pc}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/BBsema.cxl}
\end{code}
\caption{[BBsema.cxl] Bounded Buffer implementation using semaphores.}
\label{fig:boundedbuffer}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/BBsematest.cxl}
\end{code}
\caption{[BBsematest.cxl] Test program for Figure~\ref{fig:boundedbuffer}.}
\label{fig:bbtest}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/BBsemadata.cxl}
\end{code}
\caption{[BBsemadata.cxl] Test whether the bounded buffer delivers the right data in the
right order.}
\label{fig:PCsemadata}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/gpu.cxl}
\end{code}
\caption{[gpu.cxl] A thread-unsafe GPU allocator.}
\label{fig:gpu}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/RWsbs.cxl}
\end{code}
\caption{[RWsbs.cxl] Reader/Writer Lock implementation using Split Binary Semaphores.}
\label{fig:RWsplitsema}
\end{figure}

@@

CCSs are easy to implement using busy waiting:
\begin{code}
\begin{Verbatim}[xleftmargin=5mm,numbers=left]
let blocked = True:
    while blocked:
        lock();
        if application-condition-holds:
            blocked = False;
        else:
            unlock():
@cr: application-code;
unlock():
\end{Verbatim}
\end{code}

@@

\begin{figure}
\begin{code}
{\small
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/RWfair.cxl}
}
\end{code}
\caption{[RWfair.cxl] Reader/Writer Lock SBS implementation addressing fairness.}
\label{fig:RWfair}
\end{figure}

@@

\begin{figure}
\begin{code}
{\small
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/RWqueue.cxl}
}
\end{code}
\caption{[RWqueue.cxl] Reader/Writer Lock SBS implementation using a queue.}
\label{fig:RWqueue}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/RWqtest.cxl}
\end{code}
\caption{[RWqtest.cxl] Test program for Figure~\ref{fig:RWqueue}.}
\label{fig:RWqtest}
\end{figure}

@@

\begin{code} 
\begin{Verbatim}[xleftmargin=5mm,numbers=left]
def wait(c):
    V(&mutex);
    P(c);
;
def signal(c):
    V(c);
    P(&mutex);
;
\end{Verbatim} 
\end{code}

@@

\begin{figure}
\begin{code}
\begin{Verbatim}[xleftmargin=5mm,numbers=left]
import bag;

def Condition(lk):
    result = dict{ .lock: lk, .waiters: bagEmpty() };
;
def wait(c):
    let lk = (^c).lock, blocked = True,
                cnt = bagCount((^c).waiters, nametag()):
        bagAdd(&(^c).waiters, nametag());
        ^lk = False;
        while blocked:
            atomic:
                if (not (^lk)) and
                        (bagCount((^c).waiters, nametag()) <= cnt):
                    ^lk = True;
                    blocked = False;
                ;
            ;
        ;
    ;
;
def notify(c):
    let waiters = (^c).waiters:
        if waiters != bagEmpty():
            bagRemove(&(^c).waiters, bagChoose(waiters));
        ;
    ;
;
def notifyAll(c):
    (^c).waiters = bagEmpty();
;
\end{Verbatim}
\end{code}
\caption{Implementation of condition variables in the \texttt{synch} module.}
\label{fig:cv}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/RWcv.cxl}
\end{code}
\caption{[RWcv.cxl] Reader/Writer Lock implementation using Mesa-style condition variables.}
\label{fig:RWcv}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/Diners.cxl}
\end{code}
\caption{[Diners.cxl] Dining Philosophers.}
\label{fig:diners}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/DinersCV.cxl}
\end{code}
\caption{[DinersCV.cxl] Dining Philosophers that grab both forks at the same time.}
\label{fig:dinerscv}
\end{figure}

@@

\begin{code}
\begin{Verbatim}
while forks[left]:
    wait(&conds[left]);
;
while forks[right]:
    wait(&conds[right]);
;
\end{Verbatim}
\end{code}

@@

\vspace{1em}
\begin{code}
\begin{Verbatim}[xleftmargin=5mm,numbers=left]
    if left < right:
        lock(&forks[left]);
        lock(&(orks[right]);
    else:
        unlock(&forks[right]);
        unlock(&forks[left]);
    ;
\end{Verbatim}
\end{code}
\vspace{1em}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/DinersAvoid.cxl}
\end{code}
\caption{[DinersAvoid.cxl] Dining Philosophers that carefully avoid getting into a deadlock
scenario.}
\label{fig:dinersavoid}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/bank.cxl}
\end{code}
\caption{[bank.cxl] Bank accounts.}
\label{fig:bank}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/actor.cxl}
\end{code}
\caption{[actor.cxl] A producer/consumer actor.}
\label{fig:actor}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/actortest.cxl}
\end{code}
\caption{[actortest.cxl] Test code for producer/consumer actor.}
\label{fig:actortest}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/abp.cxl}
\end{code}
\caption{[abp.cxl] Alternating Bit Protocol.}
\label{fig:abp}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/abptest.cxl}
\end{code}
\caption{[abptest.cxl] Test code for alternating bit protocol.}
\label{fig:abptest}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/hw.cxl}
\end{code}
\caption{[hw.cxl] Non-blocking queue.}
\label{fig:hw}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/lst1.cxl}
\end{code}
\caption{[lst1.cxl] List with non-blocking read operations, part 1.}
\label{fig:lst1}
\end{figure}

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/lst2.cxl}
\end{code}
\caption{[lst2.cxl] List with non-blocking read operations, part 2.}
\label{fig:lst2}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/barrier.cxl}
\end{code}
\caption{[barrier.cxl] Barrier synchronization with semaphores.}
\label{fig:barrier}
\end{figure}

>>>>

\documentclass{report}
\usepackage{fullpage}
\usepackage{imakeidx}\makeindex
\usepackage{hyperref}
\usepackage[nottoc,notlot,notlof]{tocbibind}
\usepackage[toc]{glossaries}\makeglossaries
\usepackage{graphicx}
\usepackage{framed}
\usepackage{tcolorbox}
\usepackage{verbatim}
\usepackage{fancyvrb}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{geometry}
\usepackage{enumitem}
\usepackage{etoolbox}
\usepackage{bold-extra}

@@

\newcommand{\cxlsource}[1]{
\begin{tabbing}
XX\=XXX\=XXX\kill
    \input{code/#1.tex}
\end{tabbing}
}

@@

\begin{figure}
\begin{code}
\cxlsource{triangle}
\end{code}
\caption{[triangle.cxl] Computing triangle numbers.}
\label{fig:triangle}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{Up}
\end{code}
\caption{[Up.cxl] Incrementing twice in parallel.}
\label{fig:inc}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{UpEnter}
\end{code}
\caption{[UpEnter.cxl] Broken attempt at fixing the code of Figure~\ref{fig:inc}.}
\label{fig:incenter}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{csbarebones}
\end{code}
\caption{[csbarebones.cxl] A barebones critical section with two processes.}
\label{fig:csbarebones}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{cs}
\end{code}
\caption{[cx.cxl] CXL model of a critical section.}
\label{fig:cs}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{naiveLock}
\end{code}
\caption{[naiveLock.cxl] Na\"{\i}ve implementation of a shared lock.}
\label{fig:uplock}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{naiveFlags}
\end{code}
\caption{[naiveFlags.cxl] Na\"{\i}ve use of flags to solve mutual exclusion.}
\label{fig:upflags}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{naiveTurn}
\end{code}
\caption{[naiveTurn.cxl] Na\"{\i}ve use of turn variable to solve mutual exclusion.}
\label{fig:upturn}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{Peterson}
\end{code}
\caption{[Peterson.cxl] Peterson's Algorithm}
\label{fig:peterson}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{PetersonInductive}
\end{code}
\caption{[PetersonInductive.cxl] Peterson's Algorithm with Inductive Invariant}
\label{fig:petersonproof}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{csonebit}
\end{code}
\caption{[csonebit.cxl] Mutual exclusion using a flag per process.}
\label{fig:csonebit}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{PetersonMethod}
\end{code}
\caption{[PetersonMethod.cxl] Peterson's Algorithm accessed through methods.}
\label{fig:petersonmethods}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{spinlock}
\end{code}
\caption{[spinlock.cxl] Mutual Exclusion using a ``spinlock'' based on test-and-set.}
\label{fig:tas}
\end{figure}

@@

While the CVM does not have any specific built-in interlock instructions,
it does have support for executing multiple instructions atomically.
This feature is available in the CXL language in two ways.
First, any CXL statement can be made atomic by placing a label in front
of it.  Second, a group of CXL statements can be made atomic
through its \texttt{atomic}
\index{atomic statement}
statement.
We can use \texttt{atomic} statement blocks to implement a wide variety of
interlock operations.
For example, we could fix the program in Figure~\ref{fig:inc} by
constructing an atomic increment operation for a counter, like so:
\begin{code}
\cxlsource{atomicinc}
\end{code}

@@

\begin{figure}
\begin{code}
\cxlsource{spinlockInv}
\end{code}
\caption{[spinlockInv.cxl] Checking invariants.}
\label{fig:tasinv}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{csTAS}
\end{code}
\caption{[csTAS.cxl] Fixed version of Figure~\ref{fig:uplock} using test-and-set.}
\label{fig:tas2}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{lockintf}
\end{code}
\caption{The \texttt{Lock} interface in the \texttt{synch} module.}
\label{fig:spinlocks}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{UpLock}
\end{code}
\caption{[UpLock.cxl] Program of Figure~\ref{fig:inc} fixed with a lock.}
\label{fig:incfixed}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{locksusp}
\end{code}
\caption{The \texttt{Lock} interface in the \texttt{synchS} module uses suspension.}
\label{fig:suspension}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{RW}
\end{code}
\caption{[RW.cxl] Busy-Waiting Reader/Writer Lock implementation.}
\label{fig:rwbusy}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{RWtest}
\end{code}
\caption{[RWtest.cxl] Test code for Figure~\ref{fig:rwbusy}.}
\label{fig:rwbusytest}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{RWbusychk}
\end{code}
\caption{[RWbusychk.cxl] Checking for Busy Waiting.}
\label{fig:rwblock}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{RWlock}
\end{code}
\caption{[RWlock.cxl] Reader/Writer with Two Locks.}
\label{fig:rw2lock}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{RWmulti}
\end{code}
\caption{[RWmulti.cxl] Checking that multiple readers can acquire the read lock.}
\label{fig:rwmulti}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{semaphore}
\end{code}
\caption{The \texttt{Semaphore} interface in the \texttt{synch} module.}
\label{fig:semaphore}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{DinersSema}
\end{code}
\caption{[DinersSema.cxl] Italian Philosophers.}
\label{fig:dinerssema}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{qsort}
\end{code}
\caption{[qsort.cxl] Iterative qsort() implementation.}
\label{fig:qsort}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{qsorttest}
\end{code}
\caption{[qsorttest.cxl] Test program for Figure~\ref{fig:qsort}.}
\label{fig:qsorttest}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=3in]{figures/pc-crop.pdf}
\end{center}
\caption{High-level state diagram specification of producers/consumers.}
\label{fig:pc}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{BBsema}
\end{code}
\caption{[BBsema.cxl] Bounded Buffer implementation using semaphores.}
\label{fig:boundedbuffer}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{BBsematest}
\end{code}
\caption{[BBsematest.cxl] Test program for Figure~\ref{fig:boundedbuffer}.}
\label{fig:bbtest}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{BBsemadata}
\end{code}
\caption{[BBsemadata.cxl] Test whether the bounded buffer delivers the right data in the
right order.}
\label{fig:PCsemadata}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{gpu}
\end{code}
\caption{[gpu.cxl] A thread-unsafe GPU allocator.}
\label{fig:gpu}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{RWsbs}
\end{code}
\caption{[RWsbs.cxl] Reader/Writer Lock implementation using Split Binary Semaphores.}
\label{fig:RWsplitsema}
\end{figure}

@@

CCSs are easy to implement using busy waiting:
\begin{code}
\cxlsource{busywait}
\end{code}

@@

\begin{figure}
\begin{code}
{\small
\cxlsource{RWfair}
}
\end{code}
\caption{[RWfair.cxl] Reader/Writer Lock SBS implementation addressing fairness.}
\label{fig:RWfair}
\end{figure}

@@

\begin{figure}
\begin{code}
{\small
\cxlsource{RWqueue}
}
\end{code}
\caption{[RWqueue.cxl] Reader/Writer Lock SBS implementation using a queue.}
\label{fig:RWqueue}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{RWqtest}
\end{code}
\caption{[RWqtest.cxl] Test program for Figure~\ref{fig:RWqueue}.}
\label{fig:RWqtest}
\end{figure}

@@

\begin{code}
\cxlsource{hoaremon}
\end{code}

@@

\begin{figure}
\cxlsource{mesa}
\caption{Implementation of condition variables in the \texttt{synch} module.}
\label{fig:cv}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{RWcv}
\end{code}
\caption{[RWcv.cxl] Reader/Writer Lock implementation using Mesa-style condition variables.}
\label{fig:RWcv}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{Diners}
\end{code}
\caption{[Diners.cxl] Dining Philosophers.}
\label{fig:diners}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{DinersCV}
\end{code}
\caption{[DinersCV.cxl] Dining Philosophers that grab both forks at the same time.}
\label{fig:dinerscv}
\end{figure}

@@

\begin{code}
\cxlsource{baddblwait}
\end{code}

@@

\vspace{1em}
\begin{code}
\cxlsource{dinersfix}
\end{code}
\vspace{1em}

@@

\begin{figure}
\begin{code}
\cxlsource{DinersAvoid}
\end{code}
\caption{[DinersAvoid.cxl] Dining Philosophers that carefully avoid getting into a deadlock
scenario.}
\label{fig:dinersavoid}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{bank}
\end{code}
\caption{[bank.cxl] Bank accounts.}
\label{fig:bank}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{actor}
\end{code}
\caption{[actor.cxl] A producer/consumer actor.}
\label{fig:actor}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{actortest}
\end{code}
\caption{[actortest.cxl] Test code for producer/consumer actor.}
\label{fig:actortest}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{abp}
\end{code}
\caption{[abp.cxl] Alternating Bit Protocol.}
\label{fig:abp}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{abptest}
\end{code}
\caption{[abptest.cxl] Test code for alternating bit protocol.}
\label{fig:abptest}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{hw}
\end{code}
\caption{[hw.cxl] Non-blocking queue.}
\label{fig:hw}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{lst1}
\end{code}
\caption{[lst1.cxl] List with non-blocking read operations, part 1.}
\label{fig:lst1}
\end{figure}

\begin{figure}
\begin{code}
\cxlsource{lst2}
\end{code}
\caption{[lst2.cxl] List with non-blocking read operations, part 2.}
\label{fig:lst2}
\end{figure}

@@

\begin{figure}
\begin{code}
\cxlsource{barrier}
\end{code}
\caption{[barrier.cxl] Barrier synchronization with semaphores.}
\label{fig:barrier}
\end{figure}
